---
title: "Supplimentary R Script - Project 2025"
author: "Sami Benkhellat PG55704"
date: "2025-06-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document serves a supplementary document to show certain transformations that were made to the data. With certain preloaded querys so that the reader can explore the results of the study in greater detail.

```{r}
#Please install the following packages if not already on your device
library(dplyr)
library(stringr)
```

Blastp data from the second run is loaded onto R since it was ran with format 6 the output file is a tsv for ease of use in R. Format 6 does not come with a header there fore this step is just to give colums a name.  
```{r}
# Your desired column names
colnames <- c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", 
              "qstart", "qend", "sstart", "send", "evalue", "bitscore")

# Read the file without using the first row as header
results_singles <- read.delim("anotation_restricted_run.tsv", header = FALSE)

# Assign the column names manually
colnames(results_singles) <- colnames

```

To extarct all assention numbers from the blastp output file that will serve as input for Uniprots ID mapper we turn to regular expressions with the stringr package. 
```{r}
all_asention <- str_extract(results_singles$sseqid, "(?<=\\|)[^|]+(?=\\|)")
```


For what follows:
og_querys --> refers to the number of queries obtained post gene prediction
uq_accenssion --> is the number of unique accenssion numbers in the blastp result
missing_accenssions --> refers to the number of original querys that did not yield an annotation in the blast run
duplicated_accenssions --> number of duplicated accenssion numbers in blastp results

```{r}
og_querys=12439
uq_asention=length(unique(all_asention))
total_asention=length(all_asention)
missing_asentions= og_querys - uq_asention
duplicated_asention <- sum(duplicated(all_asention))

```

Save to text file with all accenssion numbers that will be used in Unitprot ID mapper

|
v

Run Unirprot mapper and choose metadata

|
v

Import the generate dataframe from Uniprot ID Mapper 
```{r}
idmapping<- read.delim("idmapping.tsv")
```

Within the dataframe generated by Uniprot's ID mapper identidy which proteins are associated to mycotoxins by searching in Gene Oncology and protein name collums for a specific keyword. This will generate separate dataframes for every keyword that can be usefull to explore.
```{r}
# Define toxins of interest
toxins <- c("mycotoxin", "aflatoxin", "ochratoxin", "patulin", "citrilin", "fumonisin", "trichothecene")

# Initialize a list to store results
grouped_accession_numbers <- list()

# Loop through each toxin and filter + organize
for (toxin in toxins) {
  toxin_rows <- idmapping %>%
    filter(grepl(toxin, Gene.Ontology..GO., ignore.case = TRUE) |
           grepl(toxin, Protein.names, ignore.case = TRUE))
  
  grouped_accession_numbers[[toxin]] <- toxin_rows %>%
    select(Organism, Entry,Gene.Ontology..GO., Protein.names ) %>%
    arrange(Organism)
}

# Now you can access each group's result like:
#grouped_accession_numbers[["mycotoxin"]]
#grouped_accession_numbers[["aflatoxin"]]
#grouped_accession_numbers[["trichothecene"]]

```


Within the dataframe generated by Uniprot's ID mapper identidy which proteins are associated to heat resistance by searching in Gene Oncology and protein name collums for a specific keyword. This will generate separate dataframes for every keyword that can be usefull to explore.
```{r}
# Define toxins of interest
heats<- c( "heat", "hsp70", "hsp90" , "hsp30", "calcineurin")

# Initialize a list to store results
grouped_accession_numbers_heat <- list()

# Loop through each toxin and filter + organize
for (heat in heats) {
  heat_rows <- idmapping %>%
    filter(grepl(heat, Gene.Ontology..GO., ignore.case = TRUE) |
           grepl(heat, Protein.names, ignore.case = TRUE))
  
  grouped_accession_numbers_heat[[heat]] <- heat_rows %>%
    select(Organism, Entry, Gene.Ontology..GO., Protein.names) %>%
    arrange(Organism)
}

# Now you can access each group's result like:
#grouped_accession_numbers_heat[["heat"]]
#grouped_accession_numbers_heat[["hsp70"]]
#grouped_accession_numbers_heat[["hsp90"]]
#grouped_accession_numbers_heat[["tubulin"]]
#grouped_accession_numbers_heat[["calcineurin"]]
#grouped_accession_numbers_heat[["hsp30"]]

```



Within the dataframe generated by Uniprot's ID mapper identidy which proteins are associated to mycotoxins by searching in Gene Oncology and protein name collums for a specific keyword. However this joining all mycotoxin associated results into one singluar table.
```{r}
toxins <- c("mycotoxin", "aflatoxin", "ochratoxin", "patulin", "citrilin", "fumonisin", "trichothecene")

# Initialize a list to store results
grouped_accession_numbers <- list()

# Loop through each toxin and filter + organize
for (toxin in toxins) {
  toxin_rows <- idmapping%>%
    filter(grepl(toxin, Gene.Ontology..GO., ignore.case = TRUE) |
           grepl(toxin, Protein.names, ignore.case = TRUE)) %>%
    select(Organism, Entry, Gene.Ontology..GO., Protein.names) %>%
    mutate(keyword = toxin) %>%    
    arrange(Organism)
  
  grouped_accession_numbers[[toxin]] <- toxin_rows
}

# Combine all into a single data frame
combined_toxin_df <- bind_rows(grouped_accession_numbers)

```

Within the dataframe generated by Uniprot's ID mapper identidy which proteins are associated to heat resistance by searching in Gene Oncology and protein name collums for a specific keyword. However this joining all heat resistance associated results into one singluar table.
```{r}
# Define heat-related keywords
heats <- c("heat", "hsp70", "hsp90", "hsp30", "calcineurin")

# Initialize a list to store results
grouped_accession_numbers_heat <- list()

# Loop through each heat keyword, filter, label, and store
for (heat in heats) {
  heat_rows <- idmapping %>%
    filter(grepl(heat, Gene.Ontology..GO., ignore.case = TRUE) |
           grepl(heat, Protein.names, ignore.case = TRUE)) %>%
    select(Organism, Entry, Gene.Ontology..GO., Protein.names) %>%
    mutate(keyword = heat) %>%
    arrange(Organism)
  
  grouped_accession_numbers_heat[[heat]] <- heat_rows
}

# Combine all into a single data frame
combined_heat_df <- bind_rows(grouped_accession_numbers_heat)

```


Merge both annotation data frames into one
```{r}
combined_annotation_df <- bind_rows(combined_toxin_df, combined_heat_df)
```


Replace keyword to find all unique instances
```{r}
#Replace keyword to fin all unique instances 
unique(combined_annotation_df$Protein.names[grepl("calcineurin", combined_annotation_df$keyword, ignore.case = TRUE)])

unique(combined_annotation_df$Organism[grepl("calcineurin", combined_annotation_df$keyword, ignore.case = TRUE)])

```

Usefull to know the number of hits per organism based on broader keyword groups
```{r}
total_hits_per_organism <- combined_annotation_df %>%
group_by(Organism) %>%
summarise(total_hits = n()) %>%
arrange(desc(total_hits))
```

Usefull to know the number of hits per organism based on per keyword
```{r}
hits_per_organism_keyword <- combined_annotation_df %>%
group_by(Organism, keyword) %>%
summarise(hits = n()) %>%
arrange(desc(hits))

```

